---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r}
library(dplyr)
library(lattice)
```

## Loading and preprocessing the data
1. Load the data
```{r}
data <- read.csv("activity.csv")
```

2.Process/transform the data 
```{r}
doc <- tbl_df(na.omit(data)) 
```

## What is mean total number of steps taken per day?
1. Calculate the total number of steps taken per day
```{r}
stepsSum <- tapply(doc$steps, doc$date, sum)
stepsSum
```

2. Histogram of the total number of steps taken each day
```{r}
hist(stepsSum, main = "Total number of steps taken each day",
     xlab = "Total number of steps per day", ylab = "Frequency") 
``` 

3. Mean and median of the total number of steps taken per day
```{r}
mean1 <- mean(stepsSum, na.rm = TRUE)
mean1
```

```{r}
median1 <- median(stepsSum, na.rm = TRUE)
median1
```

## What is the average daily activity pattern?
1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
averDaily <- tapply(doc$steps, doc$interval, mean)
intervals <- unique(doc$interval)
plot(intervals, averDaily, type = "l",
     main = "Average daily activity pattern",
     xlab = "5-minute interval",
     ylab = "Average daily steps") 
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
df1 <- as.data.frame(cbind(averDaily, intervals))
maxSteps <- df1[df1$averDaily == max(df1$averDaily), ]
max <- maxSteps$intervals
max
```

## Imputing missing values
1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
naNum <- sum(is.na(data$steps))
naNum
```

2. Devise a strategy for filling in all of the missing values in the dataset. Filling in all of the missing values, and creating a new dataset 'df2' that is equal to the original dataset but with the missing data filled in.
```{r}
df2 <- data
for(i in 1:length(df2$steps)){
    if(is.na(df2$steps[i])){
        df2$steps[i] <- df1$averDaily[which(row.names(df1)== df2$interval[i])]
    }
}
```

3. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
```{r}
stepsSum2 <- tapply(df2$steps, df2$date, sum)
hist(stepsSum2,
     main = "Total number of steps taken each day",
     xlab = "Total number of steps per day",
     ylab = "Frequency") 
```

Mean total number of steps taken per day
```{r}
mean2 <- mean(stepsSum2)
mean2
```

Median total number of steps taken per day
```{r}
median2 <- median(stepsSum2)
median2
```

If the mean differs from the estimates from the first part of the assignment
```{r}
mean1 == mean2
```

If the median differs from the estimates from the first part of the assignment
```{r}
median1 == median2
```

What is the impact of imputing missing data on the estimates of the total daily number of steps?
-In this project, the missing values were filled with the 'mean' of the total number of steps taken per day, so that the 'mean' remained unchanged after imputing the missing data. However, imputing missing data changed the 'median' from the first part of the assignment.

## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
df2 <- tbl_df(df2)
df2 <- mutate(df2, Day = weekdays(as.Date(date)))
wday <- c("Monday", "Tuesday","Wednesday", "Thursday","Friday")
for(i in 1:length(df2$Day)){
    if(df2$Day[i] %in% wday){
        df2$Day[i] = "Weekdays"
    }
    else{
        df2$Day[i] = "Weekends"
    }
}
```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r}
df2d <- filter(df2, Day == "Weekdays") 
averDaily2 <- tapply(df2d$steps, df2d$interval, mean)
interval2 <- unique(df2d$interval)
df2d <- tbl_df(as.data.frame(cbind(averDaily2,interval2)))
df2d <- mutate(df2d, Day = "Weekdays")

df2e <- filter(df2, Day == "Weekends") 
averDaily2 <- tapply(df2e$steps, df2e$interval, mean)
interval2 <- unique(df2e$interval)
df2e <- tbl_df(as.data.frame(cbind(averDaily2,interval2)))
df2e <- mutate(df2e, Day = "Weekends")

df3 <- bind_rows(df2d, df2e)
xyplot(averDaily2~interval2 | df3$Day, data = df3, 
       type = "l", layout = c(1,2),
       main = "Activity patterns between weekdays and weekends",
       xlab = "5-minute interval",
       ylab = "Average number of steps") 
```



